---
title: "itns Test"
author: "Peter Baumgartner"
date: "`r Sys.Date()`"
output: 
    learnr::tutorial:
        progressive: true
        allow_skip: true
        theme: cerulean
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
library(learnr)
library(shinyjs)
library(tidyverse)
# see: https://cran.r-project.org/web/packages/shinyjs/vignettes/shinyjs-usage.html
shiny::addResourcePath("shinyjs", system.file("srcjs", package = "shinyjs"))

knitr::opts_chunk$set(
        message = T,
        error = T,
        warning = T,
        comment = '##',
        highlight = T,
        prompt = F,
        echo = F
        )


correct = "<font color='green'>"
incorrect = "<font color='red'>"
df_laptop <- read_csv("itns-01/Describe_Laptop.csv")
laptop <- deframe(df_laptop) # covert tibble to vector
```

<!-- see https://cran.r-project.org/web/packages/shinyjs/vignettes/shinyjs-usage.html -->
<script src="shinyjs/inject.js"></script>


```{r generate-my-own-theme, include=FALSE}

my_theme <- theme_light() +
        theme(plot.title = element_text(size = 10, face = "bold", hjust = 0.5))
        theme(plot.background = element_rect(color = NA, fill = NA)) +
        theme(plot.margin = margin(1, 0, 0, 0, unit = 'cm'))

        
```



<div class="alert alert-info" role="alert">
  <strong>Info:</strong> What follows is a very simple demonstration of data manipulation. Depending of the educational task you can elaborate these kind of exercises and write checker-function to test if the answer was correct and to give adequate and specific feedback. The following two exercises do not have these more complex functionality.
</div>

```{r laod-laptop-data}
# lap <- laptop$`Transcription%`
lap <- laptop
```


### Displaying data

The laptop data set is already loaded into memory under the variable name `lap`. The command to display the first 5 data is already typed in and program is run. Play around and try to find out how many data the laptop set contains.

```{r display-data, exercise=TRUE, exercise.eval=TRUE, exercise.setup="load-laptop-data"}
head(lap, 5)
```

```{r display-data-solution}
lap # just type the variable name to spit out all the data of the vector
# you could also try `head(lap, 1000)`
```

### Selecting data

R stores the laptop data internally as a vector. The next example shows you how to select a specific part of the laptop data. Try to find out how you could display the fifth to the seventh data point.

```{r select-laptop-data, exercise=TRUE, exercise.setup="load-laptop-data"}
lap[5]
```



```{r select-laptop-data-hint-1}
# Use the colon between start and end index of the vector
```

```{r select-laptop-data-hint-2}
lap[startIndex:endIndex]
```

```{r select-laptop-data-hint-3}
lap[5:7] # this is the solution
```


